### 2023-10-09 paper/paper.tex
boostrap paper

% !TeX root = paper.tex
% chktex-file 38

\documentclass[sigplan,review,screen,anonymous]{acmart}
\usepackage[utf8]{inputenc}

\setcopyright{none}

% ===== Authors =====

\author{Joseph W. Cutler}
\email{jwc@seas.upenn.edu}
\orcid{0000-0001-9399-9308} % chktex 8
\affiliation{%
  \institution{University of Pennsylvania}
  \city{Philadelphia}
  \state{Pennsylvania}
  \country{USA}
}

\author{Emina Torlak}
\email{emina@cs.washington.edu}
\orcid{} % chktex 8
\affiliation{%
  \institution{Amazon Web Services}
  \city{Seattle}
  \state{Washington}
  \country{USA}
}

\author{Michael Hicks}
\email{mwh@cs.umd.edu}
\orcid{} % chktex 8
\affiliation{%
  \institution{Amazon Web Services}
  \city{Arlington}
  \state{Virginia}
  \country{USA}
}

\renewcommand{\shortauthors}{Cutler et al.}

% Draft title
\title{Improving Verification Stability of Mechanized Metatheory}

\begin{abstract}
\end{abstract}

\keywords{}

\begin{document}
\maketitle

\begin{itemize}
    \item We
\end{itemize}

\bibliographystyle{ACM-Reference-Format}
\bibliography{refs}

% \newpage
\include{appendix}

\end{document}